Forward Modeling Galaxy Size in a Cosmological Context

### 2017-10-04 paper/galsize_paper.tex
progress on previous work section

\section{Introduction}
\label{sec:intro}
Some introduction goes here.

@@

\subsection{Progression from Backward to Forward Modeling}
\label{subsec:forwardsmodeling}

Our results give an archetypal demonstration of the natural scientific progression from backward to forward modeling. In backward modeling, some mapping is applied to observed galaxies to estimate the values of model quantities such as halo mass. In \citet{kravtsov13}, the model quantities mapped onto galaxies are $\mhalo$ and $\rvir;$ another classic example of  backward modeling uses a group- or cluster-finding algorithm to assign $\mhalo$ to observed galaxies \citep[e.g.,][]{berlind_etal06,yang_etal05a,rykoff_etal14}. Once the observed galaxies have been supplemented with model variables, then the relations of the galaxy-halo connection can be inferred, for example, by calculating quantities such as the mean stellar mass or quiescent fraction as a function of halo mass \citep[e.g.,][]{yang_etal05b,weinmann_etal06}.

In forward modeling, the direction of inference is turned around: a mapping is instead applied to the model quantities such as $\mhalo.$ In the case of {\tt Halotools}, this transforms a cosmological simulation into a synthetic galaxy catalog that can be directly compared with observations. This enables a richer quantitative study of modeling hypotheses relative to backward modeling. For example, Figure \ref{fig:clustering_ratio_upshot} shows how forward modeling allows us to exploit galaxy clustering measurements to quantitatively test whether $\mhalo$ or $\mstar$ is the statistical regulator of galaxy size. The ability disentangle coupled variables such as $\mstar$ and $\mhalo$ is just one example of this advantage of forward modeling. Another example is illustrated in Figure \ref{fig:censatsizehist}, in which we explore the role of splashback halos in setting galaxy size. In our forward modeling approach, this is entirely straightforward; in backward modeling, such an investigation would not even be possible without introducing additional modeling ingredients.

Backward modeling the galaxy-halo connection is useful for generating hypotheses and motivating functional forms. Forward modeling becomes necessary when the problem at hand becomes encumbered by multiple relevant variables, as is the case with galaxy size. Forward modeling also makes it possible to conduct rigorous Bayesian inference, which we consider to be the next natural step in the progression described here (see \S\ref{subsec:future} for further discussion).

@@

\ben
\item \citet{shankar_etal14} - SAMs vs. SDSS. Use Yang to study median sizes of central galaxies as a function of environment, as defined by Yang-determined halo mass. The data do not show any difference in the structural properties of early-type galaxies with environment, at fixed stellar mass. All hierarchical models considered in this work instead tend to predict a moderate to strong environmental dependence. 
\item \citet{huertas_company_etal13b} - No Evidence for a Dependence of the Mass-Size Relation of Early-type Galaxies on Environment in the Local Universe. The mass?size relation of ETGs in the local Universe does not significantly depend
on environment. At fixed stellar mass (or fixed central stellar mass density), galaxies
residing in clusters have similar sizes than the ones living in the field. satellites and central galaxies follow similar mass-size relations.
\item \citet{spindler_wake17} - central vs. satellite sizes. 
\bit
\item They treat $\sigma_0$ like $\mpeak$
\item Satellites are smaller than centrals of the same $\sigma_0$, but only for quiescent galaxies. 
\item no dependence of central-satellite mass difference on $\sigma_0$
\eit
\item \citet{somerville_etal17} - spin as the origin of scatter
\item \citet{cebrian_trujillo14} - NYU-VAGC find that field galaxies are slightly larger than high-density regions (defined by 2Mpc smoothing kernel)
\item \citet{vanderwel_etal14,kawamata_etal15,shibuya_etal15,huertas_company_etal13a,huang_etal17} - high redshift studies of size evolution with HST
\item \citet{shankar_etal17,yoon_kim17,huang_leauthaud_etal17} - high mass
\item \citet{huang_etal13} - multiple component profiles
\een

>>>>

\section{Introduction}
\label{sec:intro}

\ben
\item \citet{vanderwel_etal14,kawamata_etal15,shibuya_etal15,huertas_company_etal13a,huang_etal17} - high redshift studies of size evolution with HST
\item \citet{shankar_etal17,yoon_kim17,huang_leauthaud_etal17} - high mass
\item \citet{huang_etal13} - multiple component profiles
\een

@@

\subsection{Progression from Backwards to Forward Modeling}
\label{subsec:forwardsmodeling}

Our results give an archetypal demonstration of the natural scientific progression from backwards to forward modeling. In backwards modeling, some mapping is applied to observed galaxies to estimate the values of model quantities such as halo mass. In \citet{kravtsov13}, the model quantities mapped onto galaxies are $\mhalo$ and $\rvir;$ another classic example of backwards modeling uses a group- or cluster-finding algorithm to assign $\mhalo$ to observed galaxies \citep[e.g.,][]{berlind_etal06,yang_etal05a,rykoff_etal14}. Once the observed galaxies have been supplemented with model variables, then the relations of the galaxy-halo connection can be inferred, for example, by calculating quantities such as the mean stellar mass or quiescent fraction as a function of halo mass \citep[e.g.,][]{yang_etal05b,weinmann_etal06}.

In forward modeling, the direction of inference is turned around: a mapping is instead applied to the model quantities such as $\mhalo.$ In the case of {\tt Halotools}, this transforms a cosmological simulation into a synthetic galaxy catalog that can be directly compared with observations. This enables a richer quantitative study of modeling hypotheses relative to backwards modeling. For example, Figure \ref{fig:clustering_ratio_upshot} shows how forward modeling allows us to exploit galaxy clustering measurements to quantitatively test whether $\mhalo$ or $\mstar$ is the statistical regulator of galaxy size. The ability disentangle coupled variables such as $\mstar$ and $\mhalo$ is just one example of this advantage of forward modeling. Another example is illustrated in Figure \ref{fig:censatsizehist}, in which we explore the role of splashback halos in setting galaxy size. In our forward modeling approach, this is entirely straightforward; in backwards modeling, such an investigation would not even be possible without introducing additional modeling ingredients.

Backwards modeling the galaxy-halo connection is useful for generating hypotheses and motivating functional forms. Forward modeling becomes necessary when the problem at hand becomes encumbered by multiple relevant variables, as is the case with galaxy size. Forward modeling also makes it possible to conduct rigorous Bayesian inference, which we consider to be the next natural step in the progression described here (see \S\ref{subsec:future} for further discussion).

@@

Backwards modeling methods have been used extensively in the literature to gain insight into the relationship between galaxy mass, size, and environment. Broadly speaking, such studies proceed by using a galaxy group catalog to classify observed galaxies as centrals or satellites, and estimate their halo mass.  Employing such methods, several analyses of the \citet{yang_etal05b} group catalog have found that the $\mstar-\rhalf$ relation of early-type galaxies exhibits weak, if any, environmental dependence \citep{huertas_company_etal13b,shankar_etal14}.

A direct comparison to this finding is not possible because the conclusions drawn here pertain to $\mstar-$limited galaxy samples. As discussed in \citet{spindler_wake17}, it is entirely possible that size differences between centrals and satellites of the same mass can be accounted for by mutual covariance with an additional variable such as star formation rate or morphological type \citep[see also][for an explicit demonstration of this scenario]{lilly_carollo16}. In particular, suppose that centrals and satellites of the same mass have different early-type fractions as indicated in \citet{weinmann_etal06}, and that early- and late-type galaxies exhibit universal, but distinct, $\mstar-\rhalf$ relations. In such a case, centrals would be larger than satellites of the same mass, even though {\em early-type centrals} would have the same sizes as {\em early-type satellites}.

As discussed in \S\ref{subsec:future}, we are currently pursuing follow-up work in which we jointly model morphological type together with galaxy size. We consider forward modeling methods to be a requisite for progress on this issue, not only to properly handle the multi-dimensional nature of the problem, but also to rigorously treat systematic errors that plague inference based on galaxy group catalogs \citep[see][for a thorough discussion]{campbell_etal15}.

Our approach is closely aligned with the methods used in \citet{somerville_etal17}, who studied the empirical modeling features that are necessary to recover the tight scatter in the observed $\langle\rhalf\vert\mstar\rangle$ relation.

\ben
\item \citet{cebrian_trujillo14} - NYU-VAGC find that field galaxies are slightly larger than high-density regions (defined by 2Mpc smoothing kernel)
\een
